---
title: "R Notebook"
output: html_notebook
---
This R notebook explores the development and validation of a simple linear regression model.  We use the eruption data for the Old Faithful geyser that is located in Yellowstone national park
Our goal to develop a simple linear regression model that predicts the time to the next eruption, that is the interval by the length of previous eruption.  We will check the model requirements to ensure that the assumptions of the simple linear model are satisfied.  We will the estimated the mean square prediction error usi,ng a holdout data set, and then compare prediction performance.

The data we will use is located in the package **alr4**.  You will need to install it if you have not previously installed it.

The R documentation for this data set is given below: **oldfaith** Old Faithful Geyser data:  

## Description 

Data on eruptions of Old Faithful Geyser, October 1980. Variables are the duration in seconds of the current eruption, and the time in minutes to the next eruption. Collected by volunteers, and supplied by the Yellowstone National Park Geologist. Data was not collected between approximately midnight and 6 AM.

### Format

### This data frame contains the following columns:

1. Duration Duration in seconds
2. Interval Time to next eruption

### Source
  R. Hutchinson
  
### References
Weisberg, S. (2014). Applied Linear Regression, 4th edition. Hoboken NJ: Wiley.

### Example
head(oldfaith) 


## Preperation


```{r}
require(tidyverse)                     # I live in the tidyverse.
require(cowplot)                       # I prefer the cowplot graphics theme.
require(alr4)                          # Load package containing data.

oldfaith                               # Verify data file.
```

Note that the Duration is measured in seconds and he interval is measured in minutes. 

## Plotting the entire sample

In any applied statistical activity the first the first thing we should do is to plot the
data.  In simple linear regression analysis primary plot that we should use scatterplot. We use this plot to assess if a linear model is reasonable. 

We use **ggplot2 ** package (which is load by tidyverse package).

```{r}

ggplot( oldfaith,
        aes( x = Duration,
             y = Interval )) +
  geom_point() +
  xlab( "Duration of prior erunption(Seconds)" ) +
  ylab( "Interval to next eruption(Minutes)" ) +
  ggtitle( "Eruption interval versus duration of last eruption" ) +
  theme_cowplot()



```

Note that there appear to two clusters in this data.  It is useful to draw density histograms with density plots for the predictor and response variable.  

I have written an R function to do this.  On my computer this is found in my Source directory.  We import this function bi using the source statement.  You can find this at 
my Github project **Rsource** at **gitHubSlinkman/Rsource**.  We import this function with the R **source** statement.

We then draw the density histogram with a smooth density estimator line using my R function **plot_distribution** function.


```{r}
source('D:/R-Projects/Rsource/plot_distribution.R')  # Import function

plot_distribution( oldfaith$Duration,
                   "Duration of eruption(Seconds)",
                   "Duration of eruption(Seconds)" )
```

It appears that the population is divided into two sub-populations. The reason for this is not obvious from the data.


